---
title: "Week 2 Lecture"
output: 
  learnr::tutorial:
    progressive: false
    language:
      en:
        button:
          runcode: Test Code
          submitanswer: Run Code
          questionsubmit: Submit Answer
          questiontryagain: Change my answer
runtime: shiny_prerendered
description: Forests of the Future Week 9 Assignment
---

```{css echo=FALSE}
@media print {
  .topicsContainer,
  .topicActions,
  .exerciseActions .skip {
    display: none;
  }
  .topics .tutorialTitle,
  .topics .section.level2,
  .topics .section.level3:not(.hide) {
    display: block;
  }
  .topics {
    width: 100%;
  }
  .tutorial-exercise, .tutorial-question {
    page-break-inside: avoid;
  }
  .section.level3.done h3 {
    padding-left: 0;
    background-image: none;
  }
  .topics .showSkip .exerciseActions::before {
    content: "Topic not yet completed...";
    font-style: italic;
  }
}
```

```{r setup_hide, include=FALSE}
library(learnr)
library(sdmpack)

custom_checker <- function(label, user_code, check_code, envir_result, evaluate_result, envir_prep, last_value, stage, ...) {
  # this is a code check
  if(stage == "check") {
    
    sdmpack::send_env_to_RStudio(envir_prep)
    
    sdmpack::set_env(envir_result)
    
    # try(save.image(file.path(system.file(package = "fofpack"), "week_5_progress.Rdata"),
    #            safe = FALSE),
    #     silent = TRUE)

    list(message = "Code Run; Results Now Available in RStudio.", correct = TRUE, type = "success", location = "append")
    
  }

}

tutorial_options(exercise.checker = custom_checker,
                 exercise.timelimit = 300)

knitr::opts_chunk$set(echo = FALSE)
```

# Week 9 Lecture Examples

## Data Visualization

### Setup

```{r setup, include=FALSE}
library(tidyverse)
library(palmerpenguins)

data("penguins")
```

The following code loads the packages you will need for this assignment, and the dataset you will be using (and clears the workspace). Don't forget to always click 'Run Code' for each code box, which runs your code in RStudio, and allows you to access the results in later code boxes.

```{r do_setup, exercise=TRUE}
rm(list = ls())
library(tidyverse)
library(palmerpenguins)

data("penguins")

penguins
```

```{r do_setup-check}
ls()
```

The `palmerpenguins` is a dataset containing various measurements from individuals of three different penguin species. The first thing we should generally do when working with a new dataset is to try and visualize it, if only to get a better sense of it. Let's make our first plot with `ggplot2`, the main visualization package in the `tidyverse`.

### Exercise 1

Here is an example of an exercise. This is just like what you will see in your regular assignments.

This is what `ggplot2` code for making a basic plot looks like:

```{r plot1, exercise=TRUE}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point()
```

```{r plot1-check}
ls()
```

`ggplot2` plots are built up in layers using the `+` operator. Each type of layer is a different function which are added together with `+` to built a full plot. The `ggplot()` function initializes the plot. Try deleting the second layer in the code box to see what you get.

```{r q1}
question_radio("What do you think about the grey background that is the default in {ggplot}?",
              answer("I love it!", correct = TRUE, message = "Beauty is in the eye of the beholder, isn't it?"),
              answer("It's hideous.", correct = TRUE, message = "You are correct."),
              answer("It doesn't matter because I can change it if I want to.", correct = TRUE,
                     message = "This is actually the really reaallly correct answer"),
              allow_retry = TRUE)

```

Using the code box above, change the overall look of the plot by adding a theme. Try adding `+ theme_linedraw()` to the end of the code. I personally prefer `theme_minimal()` for most of my plots, but you can use whatever you like. There are also many R packages that provide different themes which do not come with the `ggplot2` package by default. If you want to completely customize your plot's appearance you can do so use the `theme()` function. Try typing `?theme` to see all the option available. We will go over another way to change the appearance of your plot without having to type a novel into your script a little bit later.

### Add some more complexity

In `ggplot2` we can 'map' a 'variable' to an 'aesthetic' to distinguish different parts of the plot. Now, that is a lot of terminology in one sentence you may not have heard before.

See Lecture slides!

Let's change the color of the points to reflect the sex of the penguins.

```{r plot2, exercise=TRUE}
ggplot(penguins, aes(bill_length_mm, body_mass_g)) +
  geom_point(aes(color = sex)) +
  theme_minimal()
```

```{r plot2-check}
ls()
```

Overall, there isn't a clear separation by sex, but maybe things will look clearer if we take into account that we have three different species here. We can separate out the species by using panels (called 'facets' in `ggplot2`). This is easy to do by adding another layer to the plot.

```{r plot3, exercise=TRUE}
ggplot(penguins, aes(bill_length_mm, body_mass_g)) +
  geom_point(aes(color = sex)) +
  facet_grid(cols = vars(species)) +
  theme_minimal()
```

```{r plot3-check}
ls()
```

### Using a GUI: `esquisse`

There is an overwhelming number of different ways of plotting data in `gglot2` and sometimes it can be tedious to tweak every aspect of a plot by painstakingly changing the code. One way to help with this is to use the `ggplot2` graphical user interface provided by the package `esquisse` (meaning 'sketch' in french). This allows you to manipulate the plot with a GUI but then generate the code needed to make the plot, which gives the benefit of reproducibility.

```{r plot4, exercise=TRUE}
library(esquisse)
esquisser()
```

```{r plot4-check}
ls()
```

That actually won't work! That is because `esquisse` is what's called a shiny app, and this tutorial is already a shiny app. A shiny app can't run a shiny app. But we can run our code in 'RStudio' and then paste the code below.

```{r plot5, exercise=TRUE}








```

```{r plot5-check}
ls()
```

## Data Wrangling

But how did the neat and tidy `penguins` data set come to be. Well, it was 'wrangled' by a data scienctist from a messy raw data set. Data wrangling can be a major part of a data scientist's (or any scientist's) job. We will cover some basic data wrangling using `tidyverse` packages.

Let's take a look at the raw data (and load a few more useful packages).

```{r wrangle_penguin1, exercise=TRUE}
library(lubridate)
library(janitor)

penguins_raw

```

```{r wrangle_penguin1-check}
ls()
```

The most important concept in `tidyverse` data wrangling is the idea of a data 'verb'. You use 'verbs' to perform actions in a stepwise fashion on data.

See Lecture slides for more info.

The complete code used to get from `penguins_raw` to `penguins` is in the code box below. You don't need to know or understand all of these steps but it should give you and idea of the power of `tidyverse` verbs and the pipe for concisely wrangling data.

```{r complete_penguin, exercise=TRUE}
library(lubridate)
library(janitor)

penguins_df <- penguins_raw |>
  clean_names() |>
  mutate(species_short = word(species, 1)) |>
  mutate(sex = tolower(sex)) |>
  mutate(year = as.integer(lubridate::year(date_egg))) |>
  mutate(across(where(is.character), as.factor)) |>
  mutate(flipper_length_mm = as.integer(flipper_length_mm)) |>
  mutate(body_mass_g = as.integer(body_mass_g)) |>
  rename(bill_length_mm = culmen_length_mm,
         bill_depth_mm = culmen_depth_mm) |>
  select(species_short,
         island,
         bill_length_mm,
         bill_depth_mm,
         flipper_length_mm,
         body_mass_g,
         sex,
         year) |>
  rename(species = species_short) |>
  as.data.frame()

```

```{r complete_penguin-check}
ls()
```

### Exercise 2: Summarizing Data

A common task of data wrangling is summarizing. Summarizing means to take a dataset and make it smaller by using summary statistics over groups. Summary statistics are things like the mean, the median, the variance, the range, etc. You will often hear about, for example, 'median household income' as an indicator of the economic advantage of different areas or groups of people. This is a summary over a grouping variable. Let's try creating some summaries of our penguin data.

```{r summary_penguin, exercise=TRUE}

penguins_summ <- penguins |>
  group_by(species, sex) |> ## sets variables as 'grouping' variables for later verbs
  summarise(mean_body_mass = mean(body_mass_g),
            var_body_mass = var(body_mass_g))

penguins_summ

```

```{r summary_penguin-check}
ls()
```

## Packages used

To install all the packages used today, which you will need for this weeks assignment you can run following to check which are already installed (any packages missing will throw a warning)

```{r check_install, exercise=TRUE}
require(tidyverse)
require(palmerpenguins)
require(esquisse)
require(lubridate) # not required for assignment
require(janitor) # not required for assignment
```

```{r check_install-check}
ls()
```

Install any one you missed by copying from the code below and pasting it into the RStudio console (it won't work if you run it inside the tutorial).

```{r install, exercise=TRUE}
install.packages(tidyverse)
install.packages(palmerpenguins)
install.packages(esquisse)
install.packages(lubridate)
install.packages(janitor)
```

```{r install-check}
ls()
```

## Submit

```{r context="server"}
learnrhash::encoder_logic()
```

Once you have completed the assignment to your satisfaction, please click the 'generate' button below. This will create a text code that you can copy and paste into the assignment text submission form on canvas, and which I can use to regenerate your answers. Please make sure you copy the entire code. The easiest way is to click the copy button in at the top right. This will copy the entire code to the clipboard.  


```{r encode, echo=FALSE}
learnrhash::encoder_ui(ui_before = NULL)
```

<!--
### Decode

```{r context="server"}
learnrhash::decoder_logic()
```

```{r decode, echo=FALSE}
learnrhash::decoder_ui()
```
-->

