---
title: "FIU SDM Course: Week 7 Lecture"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
description: SDM Course Week 7 Lecture
---

```{css echo=FALSE}
@media print {
  .topicsContainer,
  .topicActions,
  .exerciseActions .skip {
    display: none;
  }
  .topics .tutorialTitle,
  .topics .section.level2,
  .topics .section.level3:not(.hide) {
    display: block;
  }
  .topics {
    width: 100%;
  }
  .tutorial-exercise, .tutorial-question {
    page-break-inside: avoid;
  }
  .section.level3.done h3 {
    padding-left: 0;
    background-image: none;
  }
  .topics .showSkip .exerciseActions::before {
    content: "Topic not yet completed...";
    font-style: italic;
  }
}
```

```{r setup_hide, include=FALSE}
library(learnr)
library(sdmpack)

custom_checker <- function(label, user_code, check_code, envir_result, evaluate_result, envir_prep, last_value, stage, ...) {
  # this is a code check
  if(stage == "check") {
    
    #fofpack::send_env_to_RStudio(envir_prep)
    
    rstudioapi::sendToConsole(user_code, focus = TRUE)
    
    sdmpack::set_env(envir_result)

    list(message = "Code Run; Results Now Available in RStudio.", correct = TRUE, type = "success", location = "append")
    
  }

}

tutorial_options(exercise.checker = custom_checker,
                 exercise.timelimit = 1200)

knitr::opts_chunk$set(echo = FALSE)
```

```{r setup, include=FALSE}
library(sdmpack)
library(tidyverse)
library(tidymodels)

options(datatable.alloccol = 1024,
        datatable.verbose = FALSE)

data("RF_abund")
fish_dat <- RF_abund %>%
  filter(SpeciesName == "Thalassoma pavo") %>%
  mutate(Presence = as.factor(Presence))

```

## Full Model Fitting Procedure

- Going back to our reef fish example, we will go through the full process of fitting a model

---

```{r RF2, exercise=TRUE}
library(fofpack)
library(tidyverse)
library(tidymodels)

data("RF_abund")
fish_dat <- RF_abund %>%
  filter(SpeciesName == "Thalassoma pavo") %>%
  mutate(Presence = as.factor(Presence))
```

```{r RF2-check}
ls()
```

## 1. Make a Test Set

- Split off a test set to evaluate model at the very end!

```{r RF22, exercise=TRUE}
set.seed(1234)
data_split <- initial_split(fish_dat, 0.8, strata = Presence)

train_data <- training(data_split)
test_data <- testing(data_split)

train_data
test_data
```


```{r RF22-check}
ls()
```

## 2. Resamples for Cross Validation

```{r model11, exercise=TRUE}
fish_folds <- vfold_cv(train_data, v = 6, repeats = 2, strata = Presence)
fish_folds
```

```{r model11-check}
ls()
```

## 3. Make a `recipe`

- Transform data to make it less skewed and centered on zero.

```{r recipe2, exercise=TRUE}
RF_recipe <- recipe(train_data,
                    Presence ~ MeanTemp_CoralWatch +
                      + MinTemp_CoralWatch + MaxTemp_CoralWatch + SDTemp_CoralWatch + Depth_Site) %>%
  step_YeoJohnson(MeanTemp_CoralWatch, MinTemp_CoralWatch, MaxTemp_CoralWatch, SDTemp_CoralWatch, Depth_Site) %>%
  step_normalize(MeanTemp_CoralWatch, MinTemp_CoralWatch, MaxTemp_CoralWatch, SDTemp_CoralWatch, Depth_Site)
RF_recipe
```

```{r recipe2-check}
ls()
```

## 4. Make a `model`

- Setup the model object
- Choose what hyper-parameters to tune

```{r tunable, exercise = TRUE}
tunable(boost_tree())

```

---

- We use `tune()` as a placemarker for a hyper-parameter

```{r model10, exercise=TRUE}
RF_mod <-
  boost_tree(tree_depth = tune(), mtry = tune()) %>%
  set_engine('xgboost') %>%
  set_mode('classification')

RF_mod

```

```{r model10-check}
ls()
```

## 4. A `workflow` is a `recipe` and a `model`

```{r wf1, exercise=TRUE}
RF_wf <- workflow() %>%
  add_recipe(RF_recipe) %>%
  add_model(RF_mod)

RF_wf

```

```{r wf1-check}
ls()
```

## 5. Create Hyperparameter Grid and Fit Model

```{r grid, exercise=TRUE}
tree_grid <- grid_regular(tree_depth(range = c(1, 3), trans = scales::log_trans()),
                          mtry(c(1, 5)),
                          levels = 5)
tree_grid
```

```{r grid-check}
ls()
```

## 5. Run the model on a grid of hyper-parameters

```{r model12, exercise=TRUE}
RF_res <- RF_wf %>%
  tune_grid(fish_folds,
            grid = tree_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(roc_auc))

RF_res
```

```{r model12-check}
ls()
```
  
## 6. Choose Best Model(s)?

We can see the best models using the `show_best()` function.

```{r model13, exercise=TRUE}
RF_res %>%
  show_best()
```

```{r model13-check}
ls()
```

---

We can plot all the results using `collect_metrics()` to collect our `roc_auc` values and then directly feed them into a `ggplot2` plot.

```{r model_plot, exercise=TRUE}
RF_res %>%
  collect_metrics() %>%
  mutate(mtry = factor(mtry)) %>%
  ggplot(aes(tree_depth, mean, color = mtry)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  scale_color_viridis_d(begin = .95, end = 0) +
  scale_x_log10() +
  theme_minimal()
```

```{r model_plot-check}
ls()
```

## 7-8 Final Fit and Test

```{r final_fit, exercise=TRUE}
best_tree <- RF_res %>%
  select_best("roc_auc")

final_wf <- RF_wf %>% 
  finalize_workflow(best_tree)

final_wf
```

```{r final_fit-check}
ls()
```

---

```{r final_fit2, exercise=TRUE}
final_fit <- 
  final_wf %>%
  last_fit(data_split)

final_fit

```

```{r final_fit2-check}
ls()
```

---

```{r vis1, exercise = TRUE}
final_fit %>%
  collect_metrics()

RF_final_preds <- augment(final_fit$.workflow[[1]],
                          fish_dat)

ggplot(RF_final_preds, aes(MinTemp_CoralWatch, as.numeric(Presence) - 1)) +
  geom_line(aes(y = .pred_1)) +
  geom_point() +
  ylab("Probability of Presence") +
  theme_minimal()

```

## Plot Final Predictions on All Data

```{r final_fit3, exercise=TRUE}
RF_final_preds <- augment(final_fit$.workflow[[1]],
                          fish_dat)

ggplot(RF_final_preds, aes(MeanTemp_CoralWatch, as.numeric(Presence) - 1)) +
  geom_line(aes(y = .pred_1)) +
  geom_point() +
  ylab("Probability of Presence") +
  theme_minimal()
```

```{r final_fit3-check}
ls()
```

## 9. Inference?

- How are different variables important?

Importance scores:

```{r PR_final_fit2, exercise=TRUE}
library(vip)

final_fit %>% 
  extract_fit_parsnip() %>% 
  vip()

```

```{r PR_final_fit2-check}
ls()
```


